---
title: "Validation for NGFS Task 3.1"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    toc_depth: 4
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
#library(piamValidation)
devtools::load_all(".")
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(madrat)
library(piamInterfaces)
library(quitte)
library(stringr)
library(tidyr)
library(ggpubr)


knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)
# ensure that the working directory is the project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

models <- c("GCAM",
            "MESSAGEix-GLOBIOM",
            "REMIND-MAgPIE")
```

## Data

NGFS v5.0: https://doi.org/10.5281/zenodo.13989530

ScenarioMIP: internal Snapshot from 2025-10-08

Validation configuration files:

[GCAM](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_GCAM.html)

[MESSAGE](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_MESSAGE.html)

[REMIND](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_REMIND.html)

### Data sources

#### 2020 vetting

- **Primary Energy**: Deviation to IEA and BP data.
- **Final Energy**: Deviation to IEA data.
- **Capacities**: Deviation to Ember (2025).
- **Emissions|CO2|EIP**: Deviations to CEDS (2025) 
  -> only for World due to bunkers data availability

*Note*: For 2020, the tolerance range includes both the 2020 value and the 2018-2022 average
to allow for models to include or exclude a "COVID dip". In the configuration file
this is specified by comparing to the `range` of a model and its `_smoothed` copy.

#### Near-term capacities

Tested against thresholds based on 

- **Hydro**: [IEA HSMR project pipelines](https://github.com/pik-piam/mrremind/discussions/540)
- **Nuclear**: [IAEA PRIS project pipelines](https://github.com/pik-piam/mrremind/discussions/541)
- **Solar**: [BloombergNEF projections](https://github.com/pik-piam/mrremind/discussions/542) (only World available)
- **Wind**: [GWEC Global Wind Report](https://github.com/pik-piam/mrremind/discussions/543) (only World available)

#### Near-term Carbon Capture

Tested against thresholds based on [IEA CCUS project pipelines](https://github.com/pik-piam/mrremind/discussions/544).

```{r}
scens_smip_shorten <- c(
  "SSP1 - Low Emissions" = "L - SSP1",
  "SSP2 - High Emissions" = "H - SSP2",
  "SSP2 - Medium Emissions" = "M - SSP2",
  "SSP2 - Medium-Low Emissions" = "ML - SSP2",
  "SSP2 - Low Emissions" = "L - SSP2",
  "SSP2 - Low Overshoot" = "VLHO - SSP2",
  "SSP2 - Very Low Emissions" = "VLLO - SSP2"
)

regs.short_G <- c(
  "Africa_Eastern" = "EastAfrica",
  "Africa_Northern" = "NorthAfrica",
  "Africa_Southern" = "SouthAfrica",
  "Africa_Western" = "WestAfrica",
  "Argentina" = "Argentina",
  "Australia_NZ" = "AustraliaNZ",
  "Australia_and New Zealand" = "AustraliaNZ", # !SMIP
  "Brazil" = "Brazil",
  "Canada" = "Canada",
  "Central America and Caribbean" = "CentrAmCarib",
  "Central Asia" = "CentralAsia",
  "China" = "China",
  "Colombia" = "Colombia",
  "EU-12" = "EU12",
  "EU-15" = "EU15",
  "Europe_Eastern" = "EastEurope",
  "Europe_Non_EU" = "NonEUEurope",
  "European Free Trade Association" = "EFTA",
  "India" = "India",
  "Indonesia" = "Indonesia",
  "Japan" = "Japan",
  "Mexico" = "Mexico",
  "Middle East" = "MiddleEast",
  "Pakistan" = "Pakistan",
  "Russia" = "Russia",
  "South Africa" = "SouthAfrica",
  "South America_Northern" = "NouthAmerica",
  "South America_Southern" = "SouthAmericS",
  "South Asia" = "SouthAsia",
  "South Korea" = "SouthKorea",
  "Southeast Asia" = "SEAsia",
  "Taiwan" = "Taiwan",
  "USA" = "USA",
  "World" = "World"
)

regs.short_M <- c(
  "China" = "China",
  "Eastern Europe" = "EastEurope",
  "Former Soviet Union" = "FmrSovUnion",
  "Latin America and the Caribbean" = "LatAmCarib",
  "Middle East and North Africa" = "MidEastNAfr",
  "North America" = "NorthAmeric",
  "Other Pacific Asia" = "OtherPacAsi",
  "Pacific OECD" = "PacOECD",
  "Rest of Centrally Planned Asia" = "RestCPAsia",
  "Rest Centrally Planned Asia" = "RestCPAsia",
  "South Asia" = "SouthAsia",
  "Sub-Saharan Africa" = "SubSahAfric",
  "Sub-saharan Africa" = "SubSahAfric",
  "Western Europe" = "WestEurope",
  "World" = "World"
)

regs.short_R <- c(
  "Canada Australia New Zealand" = "CanAusNZ",
  "Canada, NZ, Australia" = "CanAusNZ",
  "China and Taiwan" = "China",
  "China" = "China",
  "EU 28" = "EU28",
  "India" = "India",
  "Japan" = "Japan",
  "Latin America and the Caribbean" = "LatAmCarib",
  "Middle East and North Africa" = "MidEastNAfr",
  "Middle East, North Africa, Central Asia" = "MidEastNAfr",
  "Non-EU28 Europe" = "NonEUEurope",
  "Other Asia" = "OtherAsia",
  "Russia and Reforming Economies" = "RussRefEcon",
  "Countries from the Reforming Economies of the Former Soviet Union" = "RussRefEcon",
  "Sub-Saharan Africa" = "SubSahAfric",
  "Sub-saharan Africa" = "SubSahAfric",
  "United States of America" = "USA",
  "World" = "World"
)
```


### Scenarios

```{r plot.emi.scen.letter}
cat("ScenarioMIP scenario abbreviations\n")
scens_smip_shorten
```

### Regional Abbreviations

Note that models slightly changed some region names between NGFS and ScenarioMIP.
```{r}
cat("GCAM\n")
regs.short_G
cat("MESSAGE\n")
regs.short_M
cat("REMIND\n")
regs.short_R
```


## Validation


```{r}
# Import data
# ngfs <- readxl::read_xlsx("user/Data/NGFS/phase5/1729832902777-V5.0-NGFS-Phase-5/IAM_data.xlsx") %>%
#   filter(Model != "REMIND-MAgPIE 3.3-4.8 IntegratedPhysicalDamages (median)",
#          !grepl("(R5)", Region)) %>%
#   as.quitte()
# ngfs$region <- gsub("^.*?\\|", "", ngfs$region)  # remove model name from regions
# saveRDS(ngfs, "data/ngfs.rds")
ngfs <- readRDS("data/ngfs.rds")

# scenarioMIP data after processing in "validationReport_scenarioMIP_modelNative.Rmd" 
# scen <- readRDS("data/scenario_all_2025-10-08.rds") %>%
#   filter(model %in% c("GCAM 7.1 scenarioMIP",
#                       "MESSAGEix-GLOBIOM-GAINS 2.1-M-R12",
#                       "REMIND-MAgPIE 3.5-4.11"),
#          period >= 2020)
# saveRDS(scen, "data/scen.rds")
scen <- readRDS("data/scen.rds")


# following ScenarioMIP
# TODO: still has duplicates, why?
hist <- as.quitte("data/historical_all_2025-10-14.rds") %>%
  filter(model %in% c("IRENA", "Ember", "BP", "IEA-EB-directSum", "CEDS"))
saveRDS(hist, "data/hist.rds")
hist <- readRDS("data/hist.rds")


# Harmonization ####
# model
model_matching_ngfs <- c(
  "GCAM 6.0 NGFS" = "GCAM",
  "MESSAGEix-GLOBIOM 2.0-M-R12-NGFS" = "MESSAGE",
  "REMIND-MAgPIE 3.3-4.8" = "REMIND")
model_matching_smip <- c(
  "GCAM 7.1 scenarioMIP" = "GCAM",
  "MESSAGEix-GLOBIOM-GAINS 2.1-M-R12" = "MESSAGE",
  "REMIND-MAgPIE 3.5-4.11" = "REMIND")
ngfs <- quitte::revalue.levels(ngfs, model = model_matching_ngfs)
scen <- quitte::revalue.levels(scen, model = model_matching_smip)

# variables
vars_matching <- c(
  "Carbon Sequestration|CCS" = "Carbon Capture|Geological Storage")
ngfs <- quitte::revalue.levels(ngfs, variable = vars_matching)

# scenarios
scens_ngfs_shorten <- c(
  "Nationally Determined Contributions (NDCs)" = "NDCs")

ngfs <- quitte::revalue.levels(ngfs, scenario = scens_ngfs_shorten)
scen <- quitte::revalue.levels(scen, scenario = scens_smip_shorten)

# regions
# rename ngfs regions that have different name in scenarioMIP
ngfs_to_smip_regions <- c(
  "Australia_and New Zealand" = "Australia_NZ",
  "Rest Centrally Planned Asia" = "Rest of Centrally Planned Asia",
  "Sub-saharan Africa" ="Sub-Saharan Africa",
  "Canada, NZ, Australia" = "Canada Australia New Zealand",
  "Countries from the Reforming Economies of the Former Soviet Union" =  
    "Russia and Reforming Economies",
  #"China and Taiwan" = "China",  
  #"China" = "China",
  "Middle East, North Africa, Central Asia" = "Middle East and North Africa"
) 
ngfs <- quitte::revalue.levels(ngfs, region = ngfs_to_smip_regions)
# rename China only for REMIND as it is the correct name for other models
ngfs[(ngfs$model == "REMIND" & ngfs$region == "China"), "region"] <- "China and Taiwan"


# average 2020 values over 5 year period to even out Covid shock
hist_m <- hist %>%
 filter(period %in% seq(2018, 2022)) %>%
 magclass::as.magpie(spatial = "region")
hist_m[, 2020, ] <- dimSums(hist_m, dim = 2)/5
hist_smoothed <- quitte::as.quitte(hist_m[, , ]) %>% 
  filter(period == 2020) %>%
  mutate(model = paste0(model, "_smoothed"))
hist <- rbind(hist, hist_smoothed)


# separate historical data so model names can be removed from regions without 
# creating duplicates
hist_G <- hist %>% 
  filter(grepl("GCAM|World", region)) %>%
  mutate(region = gsub("^.*?\\|", "", region))
hist_M <- hist %>% 
  filter(grepl("MESSAGE|World", region)) %>%
  mutate(region = gsub("^.*?\\|", "", region))
hist_R <- hist %>% 
  filter(grepl("REMIND|World", region)) %>%
  mutate(region = gsub("^.*?\\|", "", region))

# combine data
scen_data_all <- rbind(ngfs, scen)
data_G <- rbind(scen_data_all, hist_G)
data_M <- rbind(scen_data_all, hist_M)
data_R <- rbind(scen_data_all, hist_R)

valiData_G <- validateScenarios(data_G, "NGFS_GCAM71_2025-10-14") %>%
  filter(model == "GCAM") %>%
  appendTooltips()

valiData_M <- validateScenarios(data_M, "NGFS_MESSAGEix-GLOBIOM-GAINS21-R12_2025-10-14") %>%  
  filter(model == "MESSAGE") %>%
  appendTooltips()

valiData_R <- validateScenarios(data_R, "NGFS_REMIND-MAgPIE35-410_2025-10-14") %>%
  filter(model == "REMIND") %>%
  appendTooltips()

valiData <- rbind(valiData_G, valiData_M, valiData_R)
```


```{r}
# copy-pasted from REMIND multi-model analysis scripts
scens_ngfs <- c(
  "Delayed transition",
  "Fragmented World",
  "Current Policies",
  "NDCs",
  "Net Zero 2050",
  "Below 2°C",
  "Low demand"
)

scens_smip <- c(
  "L - SSP1",
  "H - SSP2",
  "L - SSP2",
  "VLHO - SSP2",
  "M - SSP2",
  "ML - SSP2",
  "VLLO - SSP2"
)

scen.colors <- c(
  "Delayed transition"= "#4D5CAD",
  "Fragmented World"= "#4D5CAD",
  "Current Policies"= "#4D5CAD",
  "NDCs"= "#4D5CAD",
  "Net Zero 2050"= "#4D5CAD",
  "Below 2°C"= "#4D5CAD",
  "Low demand" = "#4D5CAD",
  
  "L - SSP1"= "darkred",
  "H - SSP2"= "darkred",
  "L - SSP2"= "darkred",
  "VLHO - SSP2"= "darkred",
  "M - SSP2"= "darkred",
  "ML - SSP2"= "darkred",
  "VLLO - SSP2"= "darkred"
)

plot.model.colors <- c(
  # scenario models
  "GCAM" = "#759EA8",
  "MESSAGEix-GLOBIOM" = "#892F71",
  "REMIND-MAgPIE" = "#facb1e"
)

validationHeatmap.colors <- 
  c(green     = "#008450",
    yellow    = "#EFB700",
    red       = "#B81D13",
    cyan      = "#66ccee",
    blue      = "#4477aa",
    grey      = "#808080")

model.linetype <- c(
  "MESSAGE" = "solid",
  "GCAM" = "solid",
  "REMIND" = "solid"
)

project.linetype <-   c(
  "Delayed transition"= "dashed",
  "Fragmented World"= "dashed",
  "Current Policies"= "dashed",
  "NDCs"= "dashed",
  "Net Zero 2050"= "dashed",
  "Below 2°C"= "dashed",
  "Low demand" = "dashed",
  
  "L - SSP1"= "solid",
  "H - SSP2"= "solid",
  "L - SSP2"= "solid",
  "VLHO - SSP2"= "solid",
  "M - SSP2"= "solid",
  "ML - SSP2"= "solid",
  "VLLO - SSP2"= "solid"
)

vars.short <- c(
  "Primary Energy|Coal" = "PE|Coal",
  "Primary Energy|Gas" = "PE|Gas",
  "Primary Energy|Oil" = "PE|Oil",
  "Primary Energy|Biomass" = "PE|Biomass",
  "Emissions|CO2|Energy and Industrial Processes" = "Emi|CO2|EIP",
  "Capacity|Electricity|Nuclear" = "Cap|Nuclear",
  "Capacity|Electricity|Hydro" = "Cap|Hydro",
  "Capacity|Electricity|Solar" = "Cap|Solar",
  "Capacity|Electricity|Wind|Onshore" = "Cap|Wind|Onshore",
  "Carbon Capture|Geological Storage" = "CCS"
)
```


```{r}
# Before Plotting

# three data objects are used in the following sections:
# - valiData: contains validated data as returned by validateScenarios, 
#             needed for heat maps and line plots 
# - data: contains all relevant scenario data and meta data also outside of 
#         validation scope, needed for line plots
# - hist: contains all historical reference data, also outside of 
#         validation scope, needed for line plots

# shorten variable names
valiData <- quitte::revalue.levels(valiData, variable = vars.short)
hist <- quitte::revalue.levels(hist, variable = vars.short)
hist_G <- hist_G %>% quitte::revalue.levels(variable = vars.short)
hist_M <- hist_M %>% quitte::revalue.levels(variable = vars.short)
hist_R <- hist_R %>% quitte::revalue.levels(variable = vars.short)
scen_data_all <- quitte::revalue.levels(scen_data_all, variable = vars.short)

# shorten scenarioMIP scenario names
valiData <- quitte::revalue.levels(valiData, scenario = scens_smip_shorten) %>%
  filter(scenario %in% c(scens_ngfs, scens_smip))
scen_data_all <- quitte::revalue.levels(scen_data_all, scenario = scens_smip_shorten) %>%
  filter(scenario %in% c(scens_ngfs, scens_smip))


# shorten and harmonize region names
valiData <- valiData %>%
  quitte::revalue.levels(region = c(regs.short_G, regs.short_M, regs.short_R))
scen_data_all <- scen_data_all %>% 
  quitte::revalue.levels(region = c(regs.short_G, regs.short_M, regs.short_R))
hist_G <- hist_G %>% 
  quitte::revalue.levels(region = regs.short_G)
hist_M <- hist_M %>% 
  quitte::revalue.levels(region = regs.short_M)
hist_R <- hist_R %>% 
  quitte::revalue.levels(region = regs.short_R)


# reorder scenario names alphabetically
valiData$scenario <- factor(valiData$scenario, 
                            levels = sort(unique(valiData$scenario)))
```

```{r, fig.width=10, fig.height=6}
lineplots_world <- function(var, 
                           reg = "World",
                           vD = valiData,
                           refData = hist,
                           scenData = scen_data_all, 
                           xlim = c(2015, 2035),
                           interactive = F) {
  
p <- linePlotThresholds(valiData %>% filter(variable == var), scenData = NULL, refData = NULL, interactive = F)


  # add scenario data as lines
  if (!is.null(scenData)) {
    d <- scenData %>%
      filter(variable == var, region == reg,
             period >= min(xlim) & period <= max(xlim))
    p <- p +
      geom_line(data = d,
                aes(x = period, y = value, color = scenario, linetype = model)) +
        scale_linetype_manual(values = model.linetype) +
        scale_color_manual(values = scen.colors)
  }

  # add reference data as points
  if (!is.null(refData)) {
    h <- refData %>%
      filter(variable == var, region == reg,
             period >= min(xlim) & period <= max(xlim))
    p <- p +
      geom_point(data = h, aes(x = period, y = value, shape = model),
                 size = 1, color = "black")
  }

  if (interactive) {
    plotly::ggplotly(p) #%>%
      #layout(legend = list(title=list(text='Model, Scenario')))
  } else {
    p
  }
}
```

```{r}
# plot all regions for one model as lineplot
lineplots_model <- function(var,
                                 region = reg,
                                 model = "GCAM",
                                 vD = valiData,
                                 refData = hist_G,
                                 scenData = scen_data_all, 
                                 xlim = c(2015, 2035),
                                 interactive = F) {
t <- vD %>% filter(model == !!model, variable == var, region == !!region)
p <- linePlotThresholds(t, 
                        scenData = NULL, 
                        refData = NULL,
                        xlim = xlim,
                        interactive = F)

  # add scenario data as lines
  if (!is.null(scenData)) {
    d <- scenData %>%
      filter(variable == var, region == reg, model == !!model,
             period >= min(xlim) & period <= max(xlim))
    p <- p +
      geom_line(data = d,
                aes(x = period, y = value, color = scenario, linetype = model)) +
        scale_linetype_manual(values = model.linetype) +
        scale_color_manual(values = scen.colors) + 
      guides(fill=guide_legend(ncol=2))
  }

  # add reference data as points
  if (!is.null(refData)) {
    h <- refData %>%
      filter(variable == var, region == reg,
             period >= min(xlim) & period <= max(xlim))
    p <- p +
      geom_point(data = h, aes(x = period, y = value, shape = model),
                 size = 1, color = "black")
  }

  if (interactive) {
    plotly::ggplotly(p) #%>%
      #layout(legend = list(title=list(text='Model, Scenario')))
  } else {
    p
  }
}

```

### Overview

## Electricity 

### Cap Hydro {.tabset}

#### Heat Maps

```{r, fig.width=7, fig.height=5}
var <- "Cap|Hydro"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

#### Lineplots GCAM

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_G$region)) {
  p <- lineplots_model(var, model = "GCAM", refData = hist_G)
  print(p)
}
```

#### Lineplots MESSAGE

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_M$region)) {
  p <- lineplots_model(var, model = "MESSAGE", refData = hist_M)
  print(p)
}
```

#### Lineplots REMIND

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_R$region)) {
  p <- lineplots_model(var, model = "REMIND", refData = hist_R)
  print(p)
}
```

### Cap|Nuclear {.tabset}

#### Heat Maps

```{r, fig.width=8, fig.height=6}
var <- "Cap|Nuclear"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

#### Lineplots GCAM

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_G$region)) {
  p <- lineplots_model(var, model = "GCAM", refData = hist_G)
  print(p)
}
```

#### Lineplots MESSAGE

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_M$region)) {
  p <- lineplots_model(var, model = "MESSAGE", refData = hist_M)
  print(p)
}
```

#### Lineplots REMIND

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_R$region)) {
  p <- lineplots_model(var, model = "REMIND", refData = hist_R)
  print(p)
}
```

### Cap|Solar

```{r, fig.width=8, fig.height=6}
var <- "Cap|Solar"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

### Cap|Wind|Onshore

```{r, fig.width=8, fig.height=6}
var <- "Cap|Wind|Onshore"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```



## Energy

### PE|Coal

```{r, fig.width=8, fig.height=6}
var <- "PE|Coal"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

### PE|Gas

```{r, fig.width=8, fig.height=6}
var <- "PE|Gas"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

### PE|Oil

```{r, fig.width=8, fig.height=6}
var <- "PE|Oil"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

### PE|Biomass

Upper bound representing sustainability concerns above 100/300 EJ/yr.

```{r, fig.width=8, fig.height=6}
var <- "PE|Biomass"

xp <- "period"
yp <- "scenario"
xf <- "variable"
yf <- "region"

validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2100, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

## Carbon

### Emissions 

```{r, fig.width=8, fig.height=6}
var <- "Emi|CO2|EIP"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2020, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

### CCS {.tabset}

#### Heat Maps

```{r, fig.width=8, fig.height=6}
var <- "CCS"

xp <- "region"
yp <- "scenario"
xf <- "variable"
yf <- "period"

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "GCAM", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "MESSAGE", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)

validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_ngfs),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
validationHeatmap(
  valiData %>% 
    filter(period <= 2030, model == "REMIND", variable == var, scenario %in% scens_smip),
  main_dim = "model", interactive = F,
  x_plot = xp, y_plot = yp, x_facet = xf, y_facet = yf)
```

#### Lineplots GCAM

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_G$region)) {
  p <- lineplots_model(var, model = "GCAM", refData = hist_G)
  print(p)
}
```

#### Lineplots MESSAGE

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_M$region)) {
  p <- lineplots_model(var, model = "MESSAGE", refData = hist_M)
  print(p)
}
```

#### Lineplots REMIND

```{r, fig.width=7, fig.height=5}
for (reg in unique(hist_R$region)) {
  p <- lineplots_model(var, model = "REMIND", refData = hist_R)
  print(p)
}
```
