---
title: "NGFS Validation Phase 6 c2"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
#library(piamValidation)
devtools::load_all(".")
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(gridExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)
```

# Data
Loading runs from: `ngfs_phase_6_snapshot_C2_29_05_2026`.

Configs (not yet fully updated online):

[Scenario Boundaries Google Doc](https://docs.google.com/spreadsheets/d/1_2KVFXLJBaGYZ4VZrb-IJsyquutEvCfR/edit?gid=1397127163#gid=1397127163)

[GCAM](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_GCAM.html)

[MESSAGE](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_MESSAGE.html)

[REMIND](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_NGFS_REMIND.html)

```{r, message = FALSE, warning = FALSE}
# Data Preparation

data_dir <- "C:/Users/pascalwe/Data/NGFS/phase6/c2/ngfs_phase_6_snapshot_C2_29_05_2026/"

scen_gcm <- "ngfs_phase_6_snapshot_gcam_2026-05-21 19_00_22.807966.xlsx"
scen_msg <- "ngfs_phase_6_snapshot_message_2026-05-29 10_09_23.052160.xlsx"
scen_rmd <- "ngfs_phase_6_snapshot_remind_2026-05-29 13_57_57.070735.xlsx"

# read and save data as rds
# data_gcm <- quitte::as.quitte(paste0(data_dir, scen_gcm))
# data_msg <- quitte::as.quitte(paste0(data_dir, scen_msg))
# data_rmd <- quitte::as.quitte(paste0(data_dir, scen_rmd))

# saveRDS(rbind(data_gcm, data_msg, data_rmd), 
#   file = paste0(data_dir, "scenario_all_2026-05-29.rds"))

# # use World data of GCAM has more periods than the others, because of finer regions?
hist_gcm <- readRDS(paste0(data_dir, "historical_GCAM82_2026-06-04.rds"))
hist_msg <- readRDS(paste0(data_dir, "historical_MESSAGEix-GLOBIOM-GAINS21-R12_2026-06-04.rds"))
hist_rmd <- readRDS(paste0(data_dir, "historical_REMIND-MAgPIE36-414_2026-06-04.rds"))

hist_gcm <- hist_gcm[hist_gcm$region != "CHN", ]
hist_msg <- hist_msg[hist_msg$region != "CHN", ]
hist_rmd <- hist_rmd[hist_rmd$region != "CHN", ]

scen <- readRDS(paste0(data_dir, "scenario_all_2026-05-29.rds"))

# remove all regions from scenario data except model-native and World
scen <- scen %>%
  filter(grepl("\\||World", region)) %>%
  filter(!is.na(value)) %>%
  filter(period %in% c(seq(2020, 2060, 5), seq(2070, 2100, 10))) %>%
  mutate(region = gsub(",", "", region))  # removes "," from REMIND region CAZ

scen_gcm <- scen %>%
  filter(model == "GCAM 8.2") %>%
  mutate(region = gsub("GCAM 8.2\\|", "", region),
         model = "GCAM")
scen_msg <- scen %>%
  filter(model == "MESSAGEix-GLOBIOM-GAINS 2.1-BMT-R12") %>%
  mutate(region = gsub("MESSAGEix-GLOBIOM-GAINS 2.1-R12\\|", "", region),
         model = "MESSAGE")
scen_rmd <- scen %>%
  filter(model == "REMIND-MAgPIE 3.6-4.14") %>%
  mutate(region = gsub("REMIND-MAgPIE 3.6-4.14\\|", "", region),
         model = "REMIND")

# average 2020 values over 5 year period to even out Covid shock
hist_gcm <- average_2020(hist_gcm)
hist_msg <- average_2020(hist_msg)
hist_rmd <- average_2020(hist_rmd) %>%
  mutate(region = gsub(",", "", region))  # removes "," from REMIND region CAZ

vd_gcm <- validateScenarios(rbind(scen_gcm, hist_gcm), "NGFS_GCAM_Phase6")
vd_gcm <- appendTooltips(vd_gcm)

vd_msg <- validateScenarios(rbind(scen_msg, hist_msg), "NGFS_MESSAGE_Phase6")
vd_msg <- appendTooltips(vd_msg)

vd_rmd <- validateScenarios(rbind(scen_rmd, hist_rmd), "NGFS_REMIND_Phase6")
vd_rmd <- appendTooltips(vd_rmd)

valiData <- rbind(vd_gcm, vd_msg, vd_rmd)

# reorder regions to World is first
new_order <- unique(intersect(c("World", "GLO",
                                levels(valiData$region)), levels(valiData$region)))
new_order_gcm <- unique(intersect(c("World", "GLO",
                                levels(vd_gcm$region)), levels(vd_gcm$region)))
new_order_msg <- unique(intersect(c("World", "GLO",
                                levels(vd_msg$region)), levels(vd_msg$region)))
new_order_rmd <- unique(intersect(c("World", "GLO",
                                levels(vd_rmd$region)), levels(vd_rmd$region)))
valiData$region <- factor(valiData$region, levels = new_order)
vd_gcm$region <- factor(vd_gcm$region, levels = new_order_gcm)
vd_msg$region <- factor(vd_msg$region, levels = new_order_msg)
vd_rmd$region <- factor(vd_rmd$region, levels = new_order_rmd)

```

# Validation

```{r}
# lineplot wrapper
modelLineplots <- function(var, 
                            reg = "World",
                            vD = valiData,
                            df = data,
                            hist_models = NULL,
                            xlim = c(2005, 2035)) {

  if (reg == "all") {
    plot_list <- htmltools::tagList()
    for (i in 1:length(unique(vD$region))) {
  
      reg <- unique(vD$region)[i]
      line_data <- vD %>%
            filter(variable == var,
                   region == reg)
          
      scen_data <- df %>%
        filter(scenario != "historical",
               variable == var,
               region == reg)
          
      hist_data <- df %>%
        filter(scenario == "historical",
               variable == var,
               region == reg)
    
      if (!is.null(hist_models)) {
        hist_data <- hist_data %>% 
          filter(model %in% hist_models)
      }
      
      plot_list[[i]] <- linePlotThresholds(line_data,
                                           refData = hist_data,
                                           scenData = scen_data,
                                           xlim = xlim)
    }
    plot_list
  
  } else {
  
    line_data <- vD %>%
      filter(variable == var,
             region == reg)
    
    scen_data <- df %>%
      filter(scenario != "historical",
             variable == var,
             region == reg)
    
    hist_data <- df %>%
      filter(scenario == "historical",
             variable == var,
             region == reg)
    
    if (!is.null(hist_models)) {
      hist_data <- hist_data %>% 
        filter(model %in% hist_models)
    }
    
    linePlotThresholds(line_data,
                       refData = hist_data,
                       scenData = scen_data,
                       xlim = xlim)
    }
  }
```


## Scenario Boundaries

### Biomass {.tabset}

Global Biomass between 40 and 100 (max 120) EJ/yr for all years.

#### Heat Map

```{r, fig.width=9, fig.height=9}
var <- "Primary Energy|Biomass"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "region",
                  y_facet = "model")
```

#### Line Plot
```{r, fig.width=9, fig.height=5}
modelLineplots(var, vD = valiData, reg = "World", df = rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm), xlim = c(2010, 2100), hist_models = "IEA")
```

### PE Fossil {.tabset}

Global PE Fossil below 100 (max 150) EJ/yr for all years including and after 2060 in
net-zero scenario.

#### Heat Map

```{r, fig.width=9, fig.height=9}
var <- "Primary Energy|Fossil"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "region",
                  y_facet = "model")
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario == "o_1p5c")
modelLineplots(var, vD = valiData, reg = "World", df = d, xlim = c(2010, 2100))
```

### Carbon Capture {.tabset}

#### Heat Maps

```{r, fig.width=10, fig.height=10}
var <- "Carbon Capture|Geological Storage"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

```{r, fig.width=10, fig.height=7}
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

```{r, fig.width=10, fig.height=7}
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plot World
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm)
modelLineplots(var, vD = valiData, reg = "World", df = d, xlim = c(2010, 2100))
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = filter(vd_gcm, period < 2035), reg = "all", df = d, xlim = c(2015, 2030))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = filter(vd_msg, period < 2035), reg = "all", df = d, xlim = c(2015, 2030))
```
#### Line Plots REMIND
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = filter(vd_rmd, period < 2035), reg = "all", df = d, xlim = c(2015, 2030))
```


### Emissions CO2 {.tabset}

#### Heat Map

```{r, fig.width=9, fig.height=5}
var <- "Emissions|CO2"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "region",
                  y_facet = "model")
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario == "o_1p5c")
modelLineplots(var, vD = vd_rmd, reg = "World", df = d, xlim = c(2010, 2100))
```

### Emissions Kyoto Gases {.tabset}

#### Heat Map

```{r, fig.width=9, fig.height=5}
var <- "Emissions|Kyoto Gases"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "model",
                  y_facet = "region")
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario == "o_1p5c")
modelLineplots(var, vD = valiData, reg = "World", df = d, xlim = c(2010, 2100))
```

### Temp 50th {.tabset}

#### Heat Map

```{r, fig.width=9, fig.height=5}
var <- "AR6 climate diagnostics|Surface Temperature (GSAT)|MAGICCv7.5.3|50.0th Percentile"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "region",
                  y_facet = "model")
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario == "o_1p5c")
modelLineplots(var, vD = valiData %>% filter(scenario == "o_1p5c"), 
               reg = "World", df = d, xlim = c(2010, 2100))
```

### Temp 67th {.tabset}

#### Heat Map

```{r, fig.width=8, fig.height=5}
var <- "AR6 climate diagnostics|Surface Temperature (GSAT)|MAGICCv7.5.3|67.0th Percentile"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "scenario",
                  x_facet = "region",
                  y_facet = "model")
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario %in% c("o_1p5c"))
modelLineplots(var, vD = filter(valiData, scenario == "o_1p5c"), reg = "World", df = d, xlim = c(2010, 2100))

d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario %in% c("o_2c"))
modelLineplots(var, vD = filter(valiData, scenario == "o_2c"), reg = "World", df = d, xlim = c(2010, 2100))

d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario %in% c("d_delfrag"))
modelLineplots(var, vD = filter(valiData, scenario == "d_delfrag"), reg = "World", df = d, xlim = c(2010, 2100))

d <- rbind(scen_gcm, scen_msg, scen_rmd, hist_gcm) %>%
  filter(scenario %in% c("d_strain"))
modelLineplots(var, vD = filter(valiData, scenario == "d_strain"), reg = "World", df = d, xlim = c(2010, 2100))
```

## Historical + Near-Term Vetting

### FE  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=6}
var <- "Final Energy"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5}
data <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = vd_gcm, reg = "all", df = data,
                hist_models = c("IEA-EB-directSum", "IEA-EB-directSum_smoothed"))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5}
data <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = vd_msg, reg = "all", df = data,
                hist_models = c("IEA-EB-directSum", "IEA-EB-directSum_smoothed"))
```
#### Line Plots REMIND
```{r, fig.width=9, fig.height=5}
data <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = vd_rmd, reg = "all", df = data,
                hist_models = c("IEA-EB-directSum", "IEA-EB-directSum_smoothed"))
```

### CO2 Energy and IP  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=6}
var <- "Emissions|CO2|Energy and Industrial Processes"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

### CO2 Supply  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=6}
var <- "Emissions|CO2|Energy|Supply"
validationHeatmap(valiData[valiData$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```


## Capacities

### Nuclear  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=9}
var <- "Capacity|Electricity|Nuclear"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = vd_gcm, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = vd_msg, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots REMIND
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = vd_rmd, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```


### Hydro  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=9}
var <- "Capacity|Electricity|Hydro"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = vd_gcm, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = vd_msg, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots REMIND
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = vd_rmd, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

### Solar  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=9}
var <- "Capacity|Electricity|Solar"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = vd_gcm, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = vd_msg, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots REMIND
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = vd_rmd, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```


### Wind  {.tabset}

#### Heat Map
```{r, fig.width=9, fig.height=9}
var <- "Capacity|Electricity|Wind"
validationHeatmap(vd_gcm[vd_gcm$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_msg[vd_msg$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
validationHeatmap(vd_rmd[vd_rmd$variable == var, ],
                  x_plot = "period",
                  y_plot = "region",
                  x_facet = "scenario",
                  y_facet = "model")
```

#### Line Plots GCAM
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_gcm, hist_gcm)
modelLineplots(var, vD = vd_gcm, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots MESSAGE
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_msg, hist_msg)
modelLineplots(var, vD = vd_msg, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

#### Line Plots REMIND
```{r, fig.width=9, fig.height=5, warning=FALSE}
data <- rbind(scen_rmd, hist_rmd)
modelLineplots(var, vD = vd_rmd, reg = "all", df = data,
                hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

## Summary {.tabset}

```{r}
df_summary <- valiData

colors <- c(green     = "#008450",
            yellow    = "#EFB700",
            red       = "#B81D13",
            cyan      = "#66ccee",
            blue      = "#4477aa",
            grey      = "#808080")

# find "critical == yes" data points of each color
summary <- filter(df_summary, region == "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )


# stacked bar plot in absolute numbers of checks
p1 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")

# normalized
p2 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")


# find "critical == yes" data points of each color
summary <- filter(df_summary, region != "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )

# stacked bar plot in absolute numbers of checks
p3 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")

# normalized
p4 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")
```


### Percentage

```{r, fig.width=7, fig.height=4}
grid.arrange(p2, p4, nrow = 1)
```

### Count
```{r, fig.width=7, fig.height=4}
grid.arrange(p1, p3, nrow = 1)
```
